Fusing Filters with Integer Linear Programming

### 2014-05-05 2014betterfusionforfilters/section/03-SizeInference.tex
transducer. proof sketch.

\subsection{Subsizes and generators}
Unlike previous work, we do allow combinators with different iteration sizes to be fused together.
For example, an operation on filtered data may be fused with the filter operation that generates the data, even though the iteration sizes are different.

\begin{tabbing}
MM \= MM \= MMMMMMMMM \= MMMM \= MM \= \kill
$gen$  \>$::$\> $binds \rightarrow name \rightarrow \{name\}$ \\
$gen(bs,o)$    \\
            \> $|$ \> $o = @fold@~f~n$      \> $\in bs$ \> $=$ \> $\emptyset$ \\
            \> $|$ \> $o = @map@_n~f~ns$    \> $\in bs$ \> $=$ \> $\bigcup_{n \in ns} gen(bs, n)$ \\
            \> $|$ \> $o = @filter@~f~n$    \> $\in bs$ \> $=$ \> $\{o\}$       \\
            \> $|$ \> $o = @generate@~s~f$  \> $\in bs$ \> $=$ \> $\emptyset$ \\
            \> $|$ \> $o = @gather@~i~d$    \> $\in bs$ \> $=$ \> $gen(bs,i)$ \\
            \> $|$ \> $o = @cross@~a~b$     \> $\in bs$ \> $=$ \> $\emptyset$ \\
            \> $|$ \> $o = @external@~ins$  \> $\in bs$ \> $=$ \> $\emptyset$ \\
\end{tabbing}

%\begin{lemma}
\textbf{Lemma: unique generators}
For some bindings $bs$, then
\[
valid(\constr{bs}) \implies \forall x.\ |gen(bs, x)| \le 1
\]
%\end{lemma}

>>>>

\subsection{Transducers}
Unlike previous work, we do allow combinators with different iteration sizes to be fused together.
For example, an operation on filtered data may be fused with the filter operation that generates the data, even though the iteration sizes are different.
More generally, if the output size of a combinator is different to its iteration size, it is a \emph{transducer} from the iteration size to the output size.
As usual, a transducer may fuse with other nodes of the same iteration size,
but transducers may also fused with nodes with iteration size the same as the transducer's output size.
For our set of combinators, the only transducer is @filter@.

\begin{tabbing}
MMMM \= MM \= MMMMMMMMM \= MMMM \= MM \= \kill
$trans$  \>$::$\> $binds \rightarrow name \rightarrow \{name\}$ \\
$trans(bs,o)$    \\
            \> $|$ \> $o = @fold@~f~n$      \> $\in bs$ \> $=$ \> $\emptyset$ \\
            \> $|$ \> $o = @map@_n~f~ns$    \> $\in bs$ \> $=$ \> $\bigcup_{x \in ns} trans(bs, x)$ \\
            \> $|$ \> $o = @filter@~f~n$    \> $\in bs$ \> $=$ \> $\{o\}$       \\
            \> $|$ \> $o = @generate@~s~f$  \> $\in bs$ \> $=$ \> $\emptyset$ \\
            \> $|$ \> $o = @gather@~i~d$    \> $\in bs$ \> $=$ \> $trans(bs,i)$ \\
            \> $|$ \> $o = @cross@~a~b$     \> $\in bs$ \> $=$ \> $\emptyset$ \\
            \> $|$ \> $o = @external@~ins$  \> $\in bs$ \> $=$ \> $\emptyset$ \\
\end{tabbing}

%\begin{lemma}
\textbf{Lemma: unique transducers}.
For some bindings $bs$, each name $n$ will have at most one transducer.
\[
valid(\constr{bs}) \implies \forall n.\ |trans(bs, n)| \le 1
\]
\textbf{Proof:} by induction on $bs$. If $n = @map@_n~f~ns$,
then $trans(bs,n) = \bigcup_{x \in ns} trans(bs,x)$.
As @map@ requires its arguments to have the same size, and @filter@ introduces a fresh size for its output,
if any of the $trans(bs,x)$ are non-empty, they will refer to the same @filter@.
The other cases are trivial.
%\end{lemma}
